---
title: An introduction to the US Federal Budget with R
author: "Stephen Roecker"
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: yes
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE,  results = 'asis')
```


# Abstract

The budget of the United States Federal Government is often a contentious issue. This presentation will provide a 'rough' introduction, covering where to get the data, provide a breakdown of the numbers, and links to other resources. 


# Load packages

```{r packages}

library(dplyr)
library(ggplot2)
library(DT)
library(tidyr)

```



# Budget Data

## GitHub Snapshot

Source
- [seankross](https://github.com/seankross/usbudget)

This is an old FY15 snapshot.

```{r ghdata, eval=FALSE}
devtools::install_github("seankross/usbudget")

library(usbudget)

data("budauth")
ba <- budauth
names(ba) <- gsub("^X", "", names(ba))

```



## US Budget

Sources

- [FY19 Whitehouse OMB Data](https://www.whitehouse.gov/omb/budget)
- [Congressional Budget Office](https://www.cbo.gov/topics/budget)

```{r wh-data, eval=FALSE}

links <- list(budauth = "https://www.whitehouse.gov/wp-content/uploads/2018/02/budauth-fy2019.csv",
              outlays = "https://www.whitehouse.gov/wp-content/uploads/2018/02/outlays-fy2019.csv",
              receipts = "https://www.whitehouse.gov/wp-content/uploads/2018/02/receipts-fy2019.csv"
              )
files_raw <- lapply(links, function(x) read.csv(x, stringsAsFactors = FALSE))

# budget authority (e.g. what the federal government is allowed to spend)
files <- lapply(files_raw, function(x) {
  # remove X from fiscal years
  names(x) = gsub("^X", "", names(x))
  # convert dollars from string to numeric
  if (any(names(x) == "1962")) {
    vars = as.character(1962:2023)
    } else vars = as.character(1976:2023)
  x[vars] = lapply(x[vars], function(y) as.numeric(gsub(",", "", y)) * 1000)
  return(x)
})

budauth <- files$budauth
outlays <- files$outlays
receipts <- files$receipts

```



## USDA data

- [FY19](https://www.obpa.usda.gov/27nrcs2019notes.pdf)

```{r, eval=FALSE, echo=FALSE}
group <- "Bureau.Name"
USDA <- TRUE
top <- 10
title <- paste("Top", top, ifelse(group == "Bureau.Name", "US", "USDA"))

gg_budget <- function(group, USDA, top, title) {
  budauth2 = {budauth ->.;
    within(., {group = .[, group]}) ->.;
    if (USDA == TRUE) {
      subset(., Agency.Name == "Department of Agriculture") ->.;
      }
    vars <- as.character(1976:2023)
    .[c("group", vars)] ->.;
    reshape(., direction = "long",
            timevar = "variable", times   = vars, 
            v.names = "value",    varying = vars
            ) ->.;
    # subset(., value > 0) ->.;
    aggregate(value ~ group + variable, data = ., sum, na.rm = TRUE) ->.;
    merge(., cpi, by.x = "variable", by.y = "year", all.x = TRUE) ->.;
    within(., {
      value2 = value * adj
    }) ->.;
    }
  
  top51 = {
    subset(budauth2, variable == "2018") -> .;
    .[order(- .$value), ] ->.;
    head(., top) ->.;
    }
  
  gg <- filter(budauth2, group %in% top51$group) %>%
    ggplot(aes(x = as.numeric(variable), y = value2, col = group)) + 
    geom_line(lwd = 2) + 
    ylab("Dollars (adj for inflation)") + xlab("Year") +
    ggtitle(title) +
    scale_x_continuous(breaks = seq(1976, 2030, 8)) +
    theme(aspect.ratio = 1/2)
  
  return(gg)
  }

gg_us   <- gg_budget(group = "Agency.Name", USDA  = FALSE, top = 10, title = "Top 10 US Agencies")
gg_usda <- gg_budget(group = "Bureau.Name", USDA  = TRUE,  top = 5,  title = "Top 5 USDA Agencies")
gridExtra::grid.arrange(gg_us, gg_usda, ncol = 1)

filter(budauth2, group == "Natural Resources Conservation Service") %>%
  ggplot(aes(x = as.numeric(variable), y = value2)) +
  geom_line(lwd = 1) +
  ylab("Dollars (adj for inflation)") + xlab("Year") +
  ggtitle("NRCS") +
  scale_x_continuous(breaks = seq(1976, 2030, 8)) +
  theme(aspect.ratio = 1/2)

```



## Gross Domestic Product (GDP)

Sources

- [US Bureau of Economic Analysis](https://apps.bea.gov/iTable/iTable.cfm?ReqID=19&step=4&isuri=1&1921=flatfiles)
- [FredCast](https://fred.stlouisfed.org/)

```{r gdp, eval=FALSE}

gdp_all <- read.csv("https://apps.bea.gov/national/Release/TXT/NipaDataA.txt", stringsAsFactors = FALSE)
gdp <- subset(gdp_all, X.SeriesCode == "A001RC")[-1]
names(gdp) <- c("year", "gdp")
gdp$gdp <- as.numeric(gsub(",", "", gdp$gdp)) * 1e6

# # Data from FredCast
#gdp_raw <- read.csv("https://fred.stlouisfed.org/graph/fredgraph.csv?id=GNPA&scale=left&cosd=1929-01-01&coed=2017-01-01&fq=Annual&fam=avg&fgst=lin&fgsnd=2009-06-01&line_index=1&transformation=lin&vintage_date=2018-10-15&revision_date=2018-10-15&nd=1929-01-01", stringsAsFactors = FALSE)
# gdp$DATE <- format(as.Date(gdp$DATE), "%Y")
# names(gdp) <- c("year", "gdp")
# gdp <- subset(gdp, year %in% 1947:2018)

```



## Consumer Price Index

Sources

- [US Bureau of Labor Statistics](https://www.bls.gov/)
- [FredCast](https://fred.stlouisfed.org/)

```{r cpi_numbers, eval=FALSE}

cpi_raw <- read.csv("https://fred.stlouisfed.org/graph/fredgraph.csv?id=CPIAUCSL&cosd=1947-01-01&coed=2018-09-01&fq=Annual&fam=avg&fgst=lin&fgsnd=2009-06-01&line_index=1&transformation=lin&vintage_date=2018-10-15&revision_date=2018-10-16&nd=1947-01-01", stringsAsFactors = FALSE)
cpi <- cpi_raw
names(cpi) <- c("year", "cpi")
cpi <- within(cpi, {
  year = as.integer(format(as.Date(year), "%Y"))
  cpi  = as.numeric(cpi)
  adj = cpi[year == 2017] / cpi
  })
```


## Save Snapshot

```{r, eval=FALSE}

save(budauth, outlays, receipts, gdp, cpi, file = "C:/workspace2/usbudget.RData")

```


## Load Snapshot

```{r load-snapshot}

load(file = "C:/workspace2/usbudget.RData")

```



# Breakdown of Budget Authorization

```{r}

# FY19 Budget Authorization
formatC(sum(budauth$`2018`[budauth$`2019` > 0]), format = "fg", big.mark = ",")
formatC(sum(budauth$`2018`[budauth$`2019` < 0]), format = "fg", big.mark = ",")

# FY19 Outlays
formatC(sum(outlays$`2018`[outlays$`2019` > 0]), format = "fg", big.mark = ",")
formatC(sum(outlays$`2018`[outlays$`2019` < 0]), format = "fg", big.mark = ",")

# n_agency
length(unique(budauth$Agency.Name))

# n_bureau
length(unique(budauth$Bureau.Name))

agencies <- group_by(budauth, Agency.Name) %>%
  summarize(dol_fy18 = sum(`2018`[`2018` > 0]),
            n_bureau = length(unique(Bureau.Name)),
            n_account = length(unique(Account.Code))
            ) %>%
  arrange(- dol_fy18)

# View(agencies)

datatable(agencies)

#n_agency
length(unique(agencies$Agency.Name[agencies$n_bureau > 1]))

# Common column names
names(receipts)[names(receipts) %in% c("Agency.code", "Agency.name", "Bureau.code", "Bureau.name", "Account.code", "Account.name")] <- c("Agency.Code", "Agency.Name", "Bureau.Dode", "Bureau.Name", "Acount.Code", "Account.Name")
idx <- names(budauth) %in% names(outlays) & names(budauth) %in% names(receipts)
vars <- names(budauth)[idx]

budget <- rbind(
  data.frame(file = "budauth",  budauth[vars],  check.names = FALSE),
  data.frame(file = "outlays",  outlays[vars],  check.names = FALSE),
  data.frame(file = "receipts", receipts[vars], check.names = FALSE)
  )

select(budget, file, as.character(1976:2023)) %>%
  gather(key = year, value = dollars, - file) %>%
  group_by(file, year) %>%
  summarize(dollars = sum(dollars)) %>%
  mutate(year = as.integer(year)) %>% 
  ggplot(aes(x = year, y = dollars / 1e9, col = file)) +
  geom_line(aes(lty = (year <= 2018)), lwd = 2) +
  geom_vline(xintercept = 2018) +
  scale_linetype_manual(values = c("dotted", "solid")) +
  scale_x_continuous(breaks = seq(1880, 2030, 8), limits = c(1976, 2024)) +
  theme(aspect.ratio = 1/2) +
  ylab("Billons of Dollars") +
  ggtitle("Budget Files")

```


# Top 10 US Agencies

```{r}

bud <- select(budauth, Agency.Name, as.character(1976:2023)) %>%
  gather(key = year, value = dollars, - Agency.Name) %>%
  group_by(Agency.Name, year) %>%
  summarize(dollars = sum(dollars)) %>%
  mutate(year = as.integer(year))

top10 <- filter(bud, year == 2018) %>%
  arrange(- dollars) %>% # View()
  head()

gg_budget <- filter(bud, Agency.Name %in% top10$Agency.Name) %>%
  ggplot(aes(x = year, y = dollars / 1e9, col = Agency.Name)) +
  geom_line(aes(lty = (year <= 2018)), lwd = 2) +
  geom_vline(xintercept = 2018) +
  scale_linetype_manual(values = c("dotted", "solid")) +
  scale_x_continuous(breaks = seq(1880, 2030, 8), limits = c(1976, 2024)) +
  theme(aspect.ratio = 1/2) +
  ylab("Billons of Dollars") +
  ggtitle("Top 10 US Agencies")
plot(gg_budget)

```



# Mandatory vs Discretionary Spending

```{r man_vs_des}

sp <- select(budauth, BEA.Category, as.character(1976:2023)) %>%
  gather(key = year, value = dollars, - BEA.Category) %>%
  group_by(BEA.Category, year) %>%
  summarize(dollars = sum(dollars[dollars > 0])) %>%
  mutate(year = as.integer(year))

ggplot(sp, aes(x = year, y = dollars / 1e9, col = BEA.Category)) +
  geom_line(aes(lty = (year <= 2018)), lwd = 2) +
  geom_vline(xintercept = 2018) +
  scale_linetype_manual(values = c("dotted", "solid")) +
  scale_x_continuous(breaks = seq(1880, 2030, 8)) +
  theme(aspect.ratio = 1/2) +
  ylab("Billons of Dollars") +
  ggtitle("Spending Category")

# percentage
group_by(budauth, BEA.Category, Agency.Name) %>%
  summarize(dollars = sum(`2018`)) %>%
  mutate(dollars = ifelse(is.na(dollars), 0, dollars)) %>%
  spread(BEA.Category, dollars) %>%
  mutate(pct_man = Mandatory / (Mandatory + Discretionary)) %>%
  arrange(- Mandatory) %>%
  head(24) %>%
  datatable()

```



# Adjust for the Consumer Price Index (CPI)

Examples

 - [stackoverflow](https://stackoverflow.com/questions/12590180/inflation-adjusted-prices-package)
 - [blscraper](https://www.datascienceriot.com//r/inflation-blscraper/)
 
```{r cpi}

gg_budget_adj <- inner_join(bud, cpi, by = "year") %>%
  mutate(dollars = dollars * adj) %>%
  filter(Agency.Name %in% top10$Agency.Name) %>%
  ggplot(aes(x = year, y = dollars / 1e9, col = Agency.Name)) +
  geom_line(lwd = 2) +
  geom_vline(xintercept = 2018) +
  scale_x_continuous(breaks = seq(1880, 2030, 8), limits = c(1976, 2024)) +
  theme(aspect.ratio = 1/2) +
  ylab("Billons of Dollars (Adj for Inflation)") +
  ggtitle("Top 10 US Agencies")

gridExtra::grid.arrange(gg_budget, gg_budget_adj, ncol = 1)

```



# Percentage of GDP

```{r pct-gdp}

gg_gdp <- gdp %>%
  inner_join(cpi, by = "year") %>%
  mutate(dollars = gdp * adj) %>%
  ggplot(aes(x = year, y = dollars / 1e9)) +
  geom_line(lwd = 1) +
  scale_x_continuous(breaks = seq(1880, 2030, 8), limits = c(1976, 2018)) +
  theme(aspect.ratio = 1/2) +
  ylab("Billons of Dollars (Adj for Inflation)") +
  ggtitle("Gross Domestic Product")

gg_gdp_pct <- bud %>%
  group_by(year) %>%
  summarize(dollars = sum(dollars)) %>%
  inner_join(gdp, by = "year") %>%
  mutate(pct_gdp = dollars / gdp) %>%
  ggplot(aes(x = year, y = pct_gdp)) +
  geom_line(lwd = 1) +
  # ylim(0, 0.5) +
  scale_x_continuous(breaks = seq(1880, 2030, 8), limits = c(1976, 2018)) +
  theme(aspect.ratio = 1/2) +
  ylab("percent of GDP (%)") +
  ggtitle("US Budget")

gridExtra::grid.arrange(gg_gdp, gg_gdp_pct, ncol = 1)

```


# Further Resources

- [Wikipedia](https://en.wikipedia.org/wiki/United_States_federal_budget)
- [White House OMB](https://whitehouse.gov/omb/budget/)
- [Congressional Budget Office](https://www.cbo.gov/topics/budget)
- [NationalPriorities.org](https://www.nationalpriorities.org/)
- [USAspending.gov](https://datalab.usaspending.gov/)
- [New York Times](https://www.nytimes.com/topic/subject/us-federal-budget)
- [FiveThirtyEight](https://fivethirtyeight.com/tag/federal-budget/)
- [NYC Data Science Academy Blog](https://nycdatascience.com/blog/student-works/visualize-the-annual-us-federal-budget-since-1962/)

